#This program runs lms on data from SIBER routine from package SIAR
#This is 3/3 codes used for SIA and HAS data analysis

```{r, echo=FALSE}
library(dplyr)
library(ggplot2)
library(arm)
```

#Read Data File
This loads a file containing posterior distributions for six Trophic Niche Metrics (TNM): dNr, dCr, TA, CD, MNND, and SDNND, from 16 different reef sites. Data file also contains three reef variables: LFTADen (Lionfish density /100m2), HASAve (averaged score of habitat complexity), and lionfish removal treatment (binary, yes/no).

```{r}
FullCommNoLF <- read.table("data/FullCommNoLF.csv",sep=",", header=T)
```

Next code run linear model with each TNM as response variable regressed against the three predictive reef variables. It creates a bootstrapped routine drawing 1000 samples per site. 

```{r}
#Bootstrap dNr####
iter <- 1000
dNr <- data.frame(Int = rep(NA, iter), HASAve = rep(NA, iter), LFTADen = rep(NA, iter), 
                    RemovTreatR = rep(NA, iter), LFTADenRemovTreatR = rep(NA, iter))
for (i in 1:iter){
  boot <- FullCommNoLF %>% group_by(Site) %>% sample_n(size = 5, replace = T)
  moddNr <- standardize(lm(dNr ~ HASAve + LFTADen * RemovTreat, data= boot))
  dNr[i,] <- as.numeric(coef(moddNr))  
} 
```

#dNr 95% CI and visualize coefficient distributions

```{r}
#estimate CI (2.5%, 97.5%) 
quantile(dNr[,2], c(.025, 0.975))
quantile(dNr[,3], c(.025, 0.975))
quantile(dNr[,4], c(.025, 0.975))
quantile(dNr[,5], c(.025, 0.975))

#coefficient distributions
ggplot(data=dNr, aes(dNr[,2])) + geom_histogram() #HASAve
ggplot(data=dNr, aes(dNr[,3])) + geom_histogram() #LFTADen
ggplot(data=dNr, aes(dNr[,4])) + geom_histogram() #RemovTreatR
ggplot(data=dNr, aes(dNr[,5])) + geom_histogram() #LFTADen*RemovTreatR
```


```{r}
#Bootstrap dCr####
dCr <- data.frame(Int = rep(NA, iter), HASAve = rep(NA, iter), LFTADen = rep(NA, iter), 
                  RemovTreatR = rep(NA, iter), LFTADenRemovTreatR = rep(NA, iter))
for (i in 1:iter){
  boot <- FullCommNoLF %>% group_by(Site) %>% sample_n(size = 5, replace = T)
  moddCr <- standardize(lm(dCr ~ HASAve + LFTADen * RemovTreat, data= boot))
  dCr[i,] <- as.numeric(coef(moddCr))  
} 
```

#dCr 95% CI and visualize coefficient distributions

```{r}
#estimate parametric bootstrap CI (2.5%, 97.5%) 
quantile(dCr[,2], c(.025, 0.975))
quantile(dCr[,3], c(.025, 0.975))
quantile(dCr[,4], c(.025, 0.975))
quantile(dCr[,5], c(.025, 0.975))

ggplot(data=dCr, aes(dCr[,2])) + geom_histogram() #HASAve
ggplot(data=dCr, aes(dCr[,3])) + geom_histogram() #LFTADen
ggplot(data=dCr, aes(dCr[,4])) + geom_histogram() #RemovTreatR
ggplot(data=dCr, aes(dCr[,5])) + geom_histogram() #LFTADen*RemovTreatR
```


```{r}
#Bootstrap TA####
TA <- data.frame(Int = rep(NA, iter), HASAve = rep(NA, iter), LFTADen = rep(NA, iter), 
                 RemovTreatR = rep(NA, iter), LFTADenRemovTreatR = rep(NA, iter))
for (i in 1:iter){
  boot <- FullCommNoLF %>% group_by(Site) %>% sample_n(size = 5, replace = T)
  modTA <- standardize(lm(TA ~ HASAve + LFTADen * RemovTreat, data= boot))
  TA[i,] <- as.numeric(coef(modTA))  
}
```

#TA 95% CI and visualize coefficient distributions

```{r}
#estimate parametric bootstrap CI (2.5%, 97.5%) 
quantile(TA[,2], c(.025, 0.975))
quantile(TA[,3], c(.025, 0.975))
quantile(TA[,4], c(.025, 0.975))
quantile(TA[,5], c(.025, 0.975))

ggplot(data=TA, aes(TA[,2])) + geom_histogram() #HASAve
ggplot(data=TA, aes(TA[,3])) + geom_histogram() #LFTADen
ggplot(data=TA, aes(TA[,4])) + geom_histogram() #RemovTreatR
ggplot(data=TA, aes(TA[,5])) + geom_histogram() #LFTADen*RemovTreatR
```


```{r}
#Bootstrap CD####
CD <- data.frame(Int = rep(NA, iter), HASAve = rep(NA, iter), LFTADen = rep(NA, iter), 
                 RemovTreatR = rep(NA, iter), LFTADenRemovTreatR = rep(NA, iter))
for (i in 1:iter){
  boot <- FullCommNoLF %>% group_by(Site) %>% sample_n(size = 5, replace = T)
  modCD <- standardize(lm(CD ~ HASAve + LFTADen * RemovTreat, data= boot))
  CD[i,] <- as.numeric(coef(modCD))  
} 
```

#CD 95% CI and visualize coefficient distributions

```{r}
#estimate parametric bootstrap CI (2.5%, 97.5%) 
quantile(CD[,2], c(.025, 0.975))
quantile(CD[,3], c(.025, 0.975))
quantile(CD[,4], c(.025, 0.975))
quantile(CD[,5], c(.025, 0.975))

ggplot(data=CD, aes(CD[,2])) + geom_histogram() #HASAve
ggplot(data=CD, aes(CD[,3])) + geom_histogram() #LFTADen
ggplot(data=CD, aes(CD[,4])) + geom_histogram() #RemovTreatR
ggplot(data=CD, aes(CD[,5])) + geom_histogram() #LFTADen*RemovTreatR
```


```{r}
#Bootstrap MNND####
MNND <- data.frame(Int = rep(NA, iter), HASAve = rep(NA, iter), LFTADen = rep(NA, iter), 
                   RemovTreatR = rep(NA, iter), LFTADenRemovTreatR = rep(NA, iter))
for (i in 1:iter){
  boot <- FullCommNoLF %>% group_by(Site) %>% sample_n(size = 5, replace = T)
  modMNND <- standardize(lm(MNND ~ HASAve + LFTADen * RemovTreat, data= boot))
  MNND[i,] <- as.numeric(coef(modMNND))  
} 
```

#MNND 95% CI and visualize coefficient distributions

```{r}
#estimate parametric bootstrap CI (2.5%, 97.5%) 
quantile(MNND[,2], c(.025, 0.975))
quantile(MNND[,3], c(.025, 0.975))
quantile(MNND[,4], c(.025, 0.975))
quantile(MNND[,5], c(.025, 0.975))

ggplot(data=MNND, aes(MNND[,2])) + geom_histogram() #HASAve
ggplot(data=MNND, aes(MNND[,3])) + geom_histogram() #LFTADen
ggplot(data=MNND, aes(MNND[,4])) + geom_histogram() #RemovTreatR
ggplot(data=MNND, aes(MNND[,5])) + geom_histogram() #LFTADen*RemovTreatR
```


```{r}
#Bootstrap SDNND####
#prepare loop and bootstraped lm
SDNND <- data.frame(Int = rep(NA, iter), HASAve = rep(NA, iter), LFTADen = rep(NA, iter), 
                    RemovTreatR = rep(NA, iter), LFTADenRemovTreatR = rep(NA, iter))
for (i in 1:iter){
  boot <- FullCommNoLF %>% group_by(Site) %>% sample_n(size = 5, replace = T)
  modSDNND <- standardize(lm(SDNND ~ HASAve + LFTADen * RemovTreat, data= boot))
  SDNND[i,] <- as.numeric(coef(modSDNND))  
} 
```

#SDNND 95% CI and visualize coefficient distributions

```{r}
#estimate parametric bootstrap CI (2.5%, 97.5%) 
print(ConfIntHAS <- quantile(SDNND[,2], c(.025, 0.975)))
print(ConfIntLF <- quantile(SDNND[,3], c(.025, 0.975)))
print(ConfIntTreat <- quantile(SDNND[,4], c(.025, 0.975)))
print(ConfIntLFTreat <- quantile(SDNND[,5], c(.025, 0.975)))

ggplot(data=SDNND, aes(SDNND[,2])) + geom_histogram() #HASAve
ggplot(data=SDNND, aes(SDNND[,3])) + geom_histogram() #LFTADen
ggplot(data=SDNND, aes(SDNND[,4])) + geom_histogram() #RemovTreatR
ggplot(data=SDNND, aes(SDNND[,5])) + geom_histogram() #LFTADen*RemovTreatR
```

#Next chunk arranges coefficient mean and SD to create coefficient plot (it's a bit messy)

```{r, echo=FALSE}
# Put model estimates into temporary data.frames:
model1Frame <- data.frame(Variable = colnames(dNr),
                          Coefficient = t(dNr %>% summarise_each(funs(mean))),
                          SE = t(dNr %>% summarise_each(funs(sd))),
                          modelName = "Nitrogen range (dNr)")
model2Frame <- data.frame(Variable = colnames(dCr),
                          Coefficient = t(dCr %>% summarise_each(funs(mean))),
                          SE = t(dCr %>% summarise_each(funs(sd))),
                          modelName = "Carbon range (dCr)")
model3Frame <- data.frame(Variable = colnames(CD),
                          Coefficient = t(CD %>% summarise_each(funs(mean))),
                          SE = t(CD %>% summarise_each(funs(sd))),
                          modelName = "Distance to centroid (CD)")
model4Frame <- data.frame(Variable = colnames(TA),
                          Coefficient = t(TA %>% summarise_each(funs(mean))),
                          SE = t(TA %>% summarise_each(funs(sd))),
                          modelName = "Total niche area (TA)")
model5Frame <- data.frame(Variable = colnames(MNND),
                          Coefficient = t(MNND %>% summarise_each(funs(mean))),
                          SE = t(MNND %>% summarise_each(funs(sd))),
                          modelName = "Mean nearest neighbour distance (MNND)")
model6Frame <- data.frame(Variable = colnames(SDNND),
                          Coefficient = t(SDNND %>% summarise_each(funs(mean))),
                          SE = t(SDNND %>% summarise_each(funs(sd))),
                          modelName = "Standard deviation nearest neighbour distance (SDNND)")

# Combine these data.frames
allModelFrame <- data.frame(rbind(model1Frame, model2Frame, model3Frame, model4Frame,
                                  model5Frame, model6Frame))  # etc.

# Change variable names to order as expected in CI plot
allModelFrame$Variables <- rep(NA, length(allModelFrame[,1]))
for (i in 1:length(allModelFrame[,1])) {
  if(allModelFrame$Variable[i] %in% "Int") {
    allModelFrame$Variables[i] <- "ENo culling" 
  } 
  if(allModelFrame$Variable[i] %in% "RemovTreatR") {
    allModelFrame$Variables[i] <- "DCulling" 
  }
  if(allModelFrame$Variable[i] %in% "HASAve") {
    allModelFrame$Variables[i] <- "CHabitat complexity" 
  }
  if(allModelFrame$Variable[i] %in% "LFTADen") {
    allModelFrame$Variables[i] <- "BLionfish Density" 
  }
  if(allModelFrame$Variable[i] %in% "LFTADenRemovTreatR") {
    allModelFrame$Variables[i] <- "ALionfish density:Culling" 
  }
}

allModelFrameb <- allModelFrame
allModelFrame$Coefficient[1] <- 0
allModelFrame$Coefficient[6] <- 0
allModelFrame$Coefficient[11] <- 0
allModelFrame$Coefficient[16] <- 0
allModelFrame$Coefficient[21] <- 0
allModelFrame$Coefficient[26] <- 0
#allModelFrame <- filter(allModelFrame, Variable != "Int")

# Specify the width of your confidence intervals
interval1 <- -qnorm((1-0.9)/2)  # 90% multiplier
interval2 <- -qnorm((1-0.95)/2)  # 95% multiplier

# Plot
#pdf("graphs/AllCoeffs.pdf", width = 15, height = 10,)
zp1 <- ggplot(allModelFrame, aes(colour = modelName))
zp1 <- zp1 + geom_hline(yintercept = 0, colour = gray(1/2), lty = 2)
zp1 <- zp1 + geom_linerange(aes(x = Variables, ymin = Coefficient - SE*interval1,
                                ymax = Coefficient + SE*interval1),
                            lwd = 1, position = position_dodge(width = 1/2))
zp1 <- zp1 + geom_pointrange(aes(x = Variables, y = Coefficient, ymin = Coefficient - SE*interval2,
                                 ymax = Coefficient + SE*interval2), size = 0.8,
                             lwd = 1/2, position = position_dodge(width = 1/2),
                             shape = 21, fill = "WHITE")
zp1 <- zp1 + coord_flip() + theme_bw() 
zp1 <- zp1 + ggtitle("Comparing several models")
print(zp1)  # The trick to these is position_dodge().
#dev.off()
```


